---
title: "01_load"
author: "Julie Bruun Brockhoff (s204519), Cecilie Kejlberg Leth (s204584), Amalie Drud Nielsen (s204560), Chrysillis Hellemann Polhaus (s153664), Lise Lotte Eriksen (s154123)"
format: 
  html:
    embed-resources: true
editor: visual
---

### Load libraries

```{r}
library(tidyverse)
library(readr)
```

### Load data and save for later use

The data is loaded as tibbles from four different files, each containing a column with gene symbols and a column with expression levels. In case an expression level is empty, NA or N/A, it is saved as "na".

```{r}
#First load the raw data taking care of missing values 
SKOV3_MOCK <- read_tsv("data/_raw/GSM8331622_SKOV_MOCK.txt", na = c("", "NA", "N/A")) 
SKOV3_KO <- read_tsv("data/_raw/GSM8331624_SKOV_KO.txt", na = c("", "NA", "N/A")) 
OVCAR3_MOCK <- read_tsv("data/_raw/GSM8331623_OVCAR3_MOCK.txt", na = c("", "NA", "N/A")) 
OVCAR3_KO <- read_tsv("data/_raw/GSM8331625_OVCAR_KO.txt", na = c("", "NA", "N/A")) 
```

### Removing duplicates

If a gene occurs more than once, all duplicates are removed and the specific gene does not occur in the analysis. This is to remove undocumented reasons for duplicates.

```{r}
OVCAR3_KO <- OVCAR3_KO |> 
  group_by(`Gene Symbol`) |> 
  filter(n() == 1) |> 
  ungroup()

OVCAR3_MOCK <- OVCAR3_MOCK |>
  group_by(`Gene Symbol`) |>
  filter(n() == 1) |>
  ungroup()

SKOV3_KO <- SKOV3_KO |>
  group_by(`Gene Symbol`) |>
  filter(n() == 1) |>
  ungroup()

SKOV3_MOCK <- SKOV3_MOCK |>
  group_by(`Gene Symbol`) |>
  filter(n() == 1) |>
  ungroup()


# OVCAR3_KO <- OVCAR3_KO |> distinct(`Gene Symbol`, .keep_all = TRUE)
# OVCAR3_MOCK <- OVCAR3_MOCK |> distinct(`Gene Symbol`, .keep_all = TRUE)
# SKOV3_KO <- SKOV3_KO |> distinct(`Gene Symbol`, .keep_all = TRUE)
# SKOV3_MOCK <- SKOV3_MOCK |> distinct(`Gene Symbol`, .keep_all = TRUE)
```

### Joining data

The four tibbles are joined into one tibble with rows corresponding to the unique genes and columns corresponding to each cell line with and without KO: OVCAR3_MOCK, OVCAR3_KO, SKOV3_MOCK, and SKOV3_KO.

```{r}
data <- left_join(OVCAR3_MOCK,
                  OVCAR3_KO, 
                  by = join_by(`Gene Symbol`)) |> 
  left_join(SKOV3_MOCK, 
            by = join_by(`Gene Symbol`)) |> 
  left_join(SKOV3_KO, 
            by = join_by(`Gene Symbol`)) |> 
  rename(gene = `Gene Symbol`)
```

### Saving in data folder

```{r}
write_tsv(data, "data/01_dat_load.tsv")
```
