---
title: "03_augment"
author: "Julie Bruun Brockhoff (s204519), Cecilie Kejlberg Leth (s204584), Amalie Drud Nielsen (s204560), Chrysillis Hellemann Polhaus (s153664), Lise Lotte Eriksen (s154123)"
format: 
  html:
    embed-resources: true
editor: 
  markdown: 
    wrap: 72
editor_options: 
  chunk_output_type: inline
---

### Load packages

```{r}
library(tidyverse)
library(readr)
library(here)
```

### Load data

```{r}
data <- read_tsv(file = here("data/02_dat_clean.tsv"))
```

### Fold change calculation

The log2 fold change (log2FC) is calculated for each of the cell lines
(OVCAR3 and SKOV3) and added as the two columns log2FC_OVCAR3 and
log2FC_SKOV3 to assess the affect of HERV-K KO.

```{r}
aug_data <- data |>
  mutate(log2FC_OVCAR3 = log2((OVCAR3_KO+1) / (OVCAR3_MOCK+1)),
         log2FC_SKOV3 = log2((SKOV3_KO+1) / (SKOV3_MOCK+1))) |>
  relocate(log2FC_OVCAR3, 
           .after = OVCAR3_KO) |> 
  relocate(log2FC_SKOV3, 
           .after = SKOV3_KO)
```

### Assign regulation pattern to data

Two new columns are added (DEG_OVCAR3 and DEG_SKOV3) to the tibble that
identifies whether each gene is up- or down-regulated in each cell line.
This is based on the log2 fold change, if it is above 2 the genes are
"Upregulated", and below -2 they are "Downregulated", if they are
between -2 and 2 they are "Neutral".

```{r}
aug_data <- aug_data |> 
  mutate(DEG_OVCAR3 = case_when(log2FC_OVCAR3 >= 2 ~ "Upregulated",
                                log2FC_OVCAR3 <= -2 ~ "Downregulated",
                                log2FC_OVCAR3 < 2 & log2FC_OVCAR3 > -2 ~ "Neutral"),
         DEG_SKOV3 = case_when(log2FC_SKOV3 >= 2 ~ "Upregulated",
                               log2FC_SKOV3 <= -2 ~ "Downregulated",
                               log2FC_SKOV3 < 2 & log2FC_SKOV3 > -2 ~ "Neutral")) |>
  relocate(DEG_OVCAR3, 
           .after = log2FC_OVCAR3) |>
  relocate(DEG_SKOV3, 
           .after = log2FC_SKOV3)
```

### Save data for later use

```{r}
write_tsv(aug_data, here("data", "03_dat_aug.tsv"))
```
