---
title: "04_describe"
author: "Julie Bruun Brockhoff (s204519), Cecilie Kejlberg Leth (s204584), Amalie Drud Nielsen (s204560), Chrysillis Hellemann Polhaus (s153664), Lise Lotte Eriksen (s154123)"
format: 
  html:
    embed-resources: true
editor: visual
---

### Load packages

```{r}
library(tidyverse)
```

### Load data

```{r}
data <- read_tsv("../data/03_dat_aug.tsv")
```

### Basic info

```{r}
ncol(data)
nrow(data)
```

There are 28040 observations (genes).

There are 9 variables (MOCK, KO, log2FC and DEG (up/downregulated) for each cell line and a column with gene symbols).

### Upregulated/downregulated genes

The number of up- and downregulated genes are counted.

```{r}
data |> count(DEG_OVCAR3)
data |> count(DEG_SKOV3)
```

For OVCAR3 there are 9.834 downregulated genes, 2.568 neutral genes and 15.638 upregulated genes

For SKOV3 there are 18.945 downregulated genes, 5.115 neutral genes and 3.980 upregulated

```{r}
data |> 
  pivot_longer(
    cols = c(DEG_OVCAR3, 
             DEG_SKOV3), 
    names_to = "cell_line", 
    values_to = "expression_level"
  ) |> 
  mutate(cell_line = case_when(cell_line == "DEG_OVCAR3" ~ "OVCAR3",
                               cell_line == "DEG_SKOV3" ~ "SKOV3")
         ) |> 
  select(gene, 
         cell_line, 
         expression_level) |> 
  count(expression_level, 
        cell_line) |> 
  ggplot(mapping = aes(x = expression_level, 
                       y = n, 
                       fill = cell_line)) +
  geom_bar(stat = "identity", 
           position = "dodge") +
  geom_text(mapping = aes(label = n), 
            position = position_dodge(width = 0.9),
            vjust = 1.5) +
  theme_minimal() + 
  theme(legend.position = "bottom") +
  labs(x = "Expression level",
       y = "Count",
       title = "Distribution of DEGs in the two cell lines",
       fill = "Cell line")

```

### Overlapping genes

```{r}
#upregulated genes for both cell lines
data |> 
  filter(DEG_OVCAR3 == "Upregulated" &
         DEG_SKOV3 == "Upregulated") |>
  count()

#downregulated genes for both cell lines
data |> 
  filter(DEG_OVCAR3 == "Downregulated" &
         DEG_SKOV3 == "Downregulated") |>
  count()
```

There are 2.197 genes, which are upregulated for both cell lines

There are 8.116 genes, which are downregulated for both cell lines
